Model to study co-evolution of trade an culture

### 2015-03-27 article.tex
results simon

\usepackage{mathtools}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%

\makeatletter
\let\oldabs\abs
\def\abs{\@ifstar{\oldabs}{\oldabs*}}
\let\oldnorm\norm
\def\norm{\@ifstar{\oldnorm}{\oldnorm*}}
\makeatother

@@

\section{INTRODUCTION}\label{sec:intro}

@@

In the model presented a set of agents are each in possession of a fixed number of goods and associate a value to each of these goods. Depending on the question studied, the value can reflect the interest of the agent for the good or it can be the price at which the agent evaluate this good. The agents are also endowed with limited storage abilities for each good that can be produced.

@@

Each agent $i$ is defined by 2 vectors each of size $r$. The first correspond to the quantity of each good that the agent possesses: $$ quantity = (q^i_1,\cdots,q^i_r) $$

Each agent makes its own estimate on the price to attribute to a certain good. This price estimate is used for both selling and buying goods :
$$ values = (v^i_1,\cdots,p^i_r) $$

On top of these elements two processes are included : imitation and innovation. The first one reflects the possibility for agents to imitate other agents. For this, an agent can copy the value vector of another agent. The second process reflects the possibility for an agent to associate new values for the goods. For this, an agent can select new elements for its value vector independently of the values of other agents.

This scheme allow us to have a model achieving the two main goals listed in section~\ref{sec:intro}. First, this model allows us to test various cultural transmission mechanisms as well as trade mechanisms. The experiments presented in this article show examples of implementations for both categories. Moreover, the fact that only two processes are modified allow comparisons between simulations with the same units. Second, few parameters are present in the default model which makes it easier to analyse and reuse by other researchers.

In order to validate our model a first result from the literature on the evolutionary transmission is reproduced. We then illustrate the possibility to implement a trade mechanisms.

\subsection{Cultural Transmission}

In this article the cultural transmission tested is the ``random copy" such as presented in~\cite{mesoudi_random_2009}. The aim is to show that our model is indeed able to reproduce the main trends expected with such copy mechanism. Namely, we should observe a distribution known as the power law, displaying a line in the log-log graph of the proportion of variants with regards to the frequency of variant.

The imitation process is a ``random copy" meaning that with a low probability each agent pick randomly one agent among all and copy it's vector of values. The innovation process is triggered with a low probability and select new elements for the vector of values. These new elements are drawn from a uniform distribution. The imitation and innovation probabilities are presented with other parameters in table~\ref{tab:parameters}

\subsection{Trading Model}\label{sec:trade}

\paragraph*{Imitation}
To test a simple trading model, two notions are introduced in the imitation mechanism: need and score. The need is a quantity of product that each agent tries to obtain. This quantity is different for each product but the need for a product is the same for all agents:
$$ need = (n_1, \cdots, n_r) $$

The score reflects the ability for an agent to obtain the products it needs, it is formally computed as follow for agent $i$:

$$ \abs{\dfrac{q^i_j - n_j}{ \sqrt{\abs{(q^i_j*q^i_j)-(n_j*n_j)}}}} $$

\paragraph*{Trading}

@@

\paragraph*{Innovation}

@@

@@

\section{RESULTS}

>>>>

@@

\section{INTRODUCTION}

@@

\todo{find another title}
\subsection{Agent-based consumers and traders}
%tell the story of what is happening

The only mandatory elements are a set of agents each possessing a fixed number of goods and associate a value to each of these goods. The value can be for example either an interest in the good or the price associated to this good. The agents are also endowed with limited storage abilities for each good that can be produced.

@@

Each agent $i$ is defined by 4 vectors each of size $r$. The first correspond to the quantity of each good that the agent possesses: $$ quantity = (q^i_1,\cdots,q^i_n) $$

Each agent makes its own estimate on the price to attribute to a certain good. This price estimate is used for both selling and buying goods :
$$ values = (v^i_1,\cdots,p^i_n) $$

@@

@@

Finally, as mentioned before, each agent will consume a certain quantity of each good. This quantity can be a property of the agent, i.e. each agent consumes a different quantity of each product :
$$ consumption = (c^i_1, \cdots, c^i_n) $$

The elements listed above constitute the basic structure of the model. Different behaviours can be used for the agents and world depending on the assumption to test. This scheme allow us to have a model achieving the two main goals listed in section~XXX. First, this model has a simple behaviour and few parameters to analyse. Second, this model is flexible enough to allow the study of multiple hypothesis while conserving the same units that can therefore be compared between simulations.

\subsection{Validation}
\xrc{Validation is impossible for a theoretical model. Either you use verification if the goal is to check that implementation is correct, or design experiment for defining the experiments that will be used to explore the model.}

In order to validate our model mulitple experiments from the litterature in economy or evolution of culture have been selected to be reproduced. The basis of selection of these experiments is two fold. First, the experiment selected has to be simple and highlight clearly one effect. This restriction is to maintain the clarity of the paper. Second, these effects should consist the basis of our further studies toward understanding the principles of the roman economics. As a consequence the hypothesis tested will constitute valid hypothesis for the study of the economics and culture of the roman empire.

\subsubsection{Economical Trading}\label{trade}

@@

\subsubsection{Cultural Evolution}

In our model the ``cultural change'' is represented by the variations in the distribution of prices for each good. In order to validate our model we will try two simple biases: simple random copying and frequency-dependent copying as \cite{mesoudi_random_2009}. The aim is to show that these simple modifications can allow us to obvserve the emergence of different cultures, each modifying the price structure obtained. The price structure obtained will be also compare to the obtained by the the economical model explained at the previous section.

The production, consumption and trade phases described are not modified. However the price update phase is modified as follow. When the random copy behaviour is used each agent has a XXX\% chance to copy a new price from one agent selected randomly. Two frequency dependent behaviour are proposed. The first to be tested is the conformist behaviour with which each agent at each agent will have a XXXX\% chance to select the most common price for each good. The second to be tested is the rarity behaviour with which each agent at each iteration will have a XXX\% chance to select the least common beahviour.

\section{RESULTS}
\subsection{Random cultural transmission and random innovation process}
In our model, agent own a set of price they assign to a certain number of resources. Those prices could be seen as the importance, or the ``value''  that each agents believe each good bears. It could be seen as cultural traits that agents can modify or share with other agent. In the first experiment the value of those traits does not impact the agent  behavior neither the way they will modify and share it. 
\subsubsection{One trait}
Knowing that and following \cite{bentley_random_2004,bentley_specialisation_2005,mesoudi_random_2009}, we think that in our model those prices will behave exactly like any other cultural trait, and that the different version of prices for each good will be found in the population in proportion that follow a power law.

It's indeed what we observe. In the first experiment, in order to compare our result with those of the literature, we use only one price. We check for some of the same $\mu$ value used in \cite{bentley_specialisation_2005,mesoudi_random_2009} that the distribution of the number of variants are the same.

That is what we can see in the figure \ref{fig:allMutation}.

Each curves on the figures show the mean for 100 runs. In our experiment one has to take in account that agents modifies their prices or exchange their price only every ten steps. Between, they exchange goods given their own prices. Therefore we choose to run our simulation during 10000 step which correspond the 1000 time steps used by \cite{bentley_random_2004,mesoudi_random_2009}, for $\mu$ the innovation rate $\in {.004,.008,.016,.064}$ and for 250 and 500 agents.

\simon{reference to the R package and the theoreticla ref in the Rpackage?}
Those distribution exhibit power law like attribute and we can fit them. The result of the alpha fitted with are in the table \ref{tab:mualpha}. The alpha values represent the mean of the fitted alpha for each run computed using RPACKAGE with the standard deviation associated with each mean.
\begin{figure}[hbp]
	\begin{center}
		\includegraphics[width=7cm]{img/allmuRandMaxBentleyValues.pdf}
	\end{center}
	\caption{rand()\%RAND\_MAX}
	\label{fig:allMutation}
\end{figure}

\begin{table}
	\centering
	\begin{tabular}{cc}
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.541193&0.031384174\\
			0.008&1.549726&0.021953251\\
			0.016&1.568096&0.017025456\\
			0.064&1.651690&0.006538209\\

		\end{tabular}
		&
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.539710&0.022515901\\
			0.008&1.557460&0.017805810\\
			0.016&1.570036&0.004412181\\
			0.064&1.651521&0.009988488\\
		\end{tabular}\\
		(a) & (b)
	\end{tabular}
	\caption{Mean \& SD are done with 100 run for each with 250 agent for (a) and 500 for (b) }
	\label{tab:mualpha}
\end{table}

\subsubsection{Multiple traits co-evolving randomly}
\simon{ this part could be resumed to one phrase saying that nothing change and that we check it without showing any result moreover}
As our aims is to study the evolution of an economy where people exchange different goods, we have checked if the previous results are the same when the agents own not just one price for one good but a vector of price associated to each good available in the environment. If the innovation process assure that all price could be modified in totally independent way, when a cultural transmission occur agents copy the whole vector of price from another agent.  We show in the table~\ref{tab:alphagoods} the mean of $\alpha$ between each resource for different number of resource.

\begin{table}
	\centering
		\begin{tabular}{l|cccc}
			  & g=1  & g= 2 & g=3 & g=4 \\\hline
			$\alpha$&\\
			sd \\
		\end{tabular}
	\caption{alpha is the mean alpha between each good for 100 runs }
	\label{tab:alphagoods}
\end{table}

\subsection{ Economical  process}

In the previous experiment the estimate price of good have no impact on the agents choices and behaviors. In a real economical system it's not the case. In order to approximate such a process and to see how our framework could handle economical mechanism we introduce a new cultural transmission mechanism.

This time agent will not copying randomly the vector of price from one random individual but they will choose among the agent that produce the same good those that seem the most successful. In practice, a fixed proportion of the less successful agents will copy the prices of the more successful agent.

To express this idea of success we define a score that is maximum when you manage to have exactly the needed amount for each good. If you exchange too much or too few of one of the good your score will decreased. The score is computed in such a way that each good count for the same proportion in the final score (i.e.: managing to get only the right amount of a good with a high ``need'' value won't help you to increase your score).

Meanwhile time the innovation mechanism used in the experiment  we follow in the previous section seems not  an adequate economical mechanism. While in a cultural point a view an innovation could be anything new in you space innovation, it's seem little probable than a price will be change to something totally new.

That why we  introduce a new innovation mechanism which will also occur with a probability of $\mu$ but this time, instead of set a totally new price, it will modify the previous price by adding or subtract to it a small amount randomly choose between $0 .. \beta$.

In order to study the impact of the introduction of those new mechanism  we summarize it in the table~\ref{tab:exp}. 
\begin{table}
	\centering
	\begin{tabular}{l|c|c}
		& Rand & Eco \\\hline
		$\mu$ &A & B \\
		$\mu +\alpha$ & C & D \\
	\end{tabular}
	\caption{Experimental Setup}
	\label{tab:exp}
\end{table}

The following  experiments will last longer than the firsts one because this time the exchange mechanisms will have an impact on the evolution of prices in the population. All the next experiments are run during 200000 steps with a innovation rate $\mu=.004$ and $\beta=.001$.
